### 2023-09-01 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@d9e5089ca1aeb57d00b06ed661d22af1941cf001 🚀

Disease emergence is complex, and is driven not only by animal-human
contact, but also by the underlying evolutionary dynamics in viral
reservoirs.\textsuperscript{1} Although host richness is often used as a
superficial proxy for spillover risk,\textsuperscript{2--4} these
approaches oversimplify the relevant interspecific heterogeneity in
immunology, behavior, and other traits, and therefore overlook unique
host pools that allow for the rapid evolution of highly divergent
viruses.\textsuperscript{5} In the case of generalist pathogens like
betacoronaviruses, there is conceptual and empirical support to the idea
that these community-level mechanisms are even more
important,\textsuperscript{6} particularly given that cross-species
transmission may, as a rule, structure viral evolution more than
co-divergence with hosts.\textsuperscript{7} This creates a disconnect
between coevolutionary theory and most existing ecological frameworks
for mapping spillover risk.

The geographic mosaic theory of coevolution (GMTC) attempts to
explicitly connect microevolutionary dynamics to the macroecology and
biogeography of symbiotic interactions.\textsuperscript{8} The GMTC
posits that coevolutionary processes among pairs\textsuperscript{9} or
complexes\textsuperscript{10} of species are structured in space by the
rippling effects of abiotic conditions onto evolutionary mechanisms,
giving rise to fragmented systems with different ecologies over large
spatial extents.\textsuperscript{11} The GMTC predicts a spatial
fragmentation of coevolutionary dynamics under the joint action of three
processes:\textsuperscript{12} coevolutionary hot- and coldspots, which
appear when the intensity of \emph{interaction} (in terms of reciprocal
fitness consequences) varies spatially; selection mosaics, wherein the
intensity of \emph{selection} varies across space, driven by both the
biotic complexity of the community (locally diverse hosts and viruses
are more biotically complex) and the local favorability of the
environment;\textsuperscript{13} and trait remixing, which occurs when
coevolutionary dynamics change when community-level \emph{functional
traits} change through meta-community dynamics.

Here, we apply the GMTC to explore and explain the global biogeography
of betacoronaviruses, the group that includes SARS-CoV, MERS-CoV, and
SARS-CoV-2. In their bat reservoirs, coronaviruses evolve through a mix
of host jumps, recombination among disparate lineages, and, to a lesser
degree, co-divergence with their hosts---\textsuperscript{2}a mix of
mechanisms that creates a complex and nonlinear relationship between
host diversity and viral emergence. Working from a recently published
database of bat hosts of betacoronaviruses, we test whether spatial
structure in bat-betacoronavirus coevolution is identifiable at a global
scale. Aiming to explain these patterns, we develop a generalized
framework for applying the GMTC to host-virus interactions, with a
specific emphasis on the potential to create independent coevolutionary
dynamics (and therefore spatial fragmentation in risk) through
heterogeneity. We develop a trivariate risk assessment system that
connects each GMTC mechanism to a quantifiable aspect of host-virus
interactions: (i) viral sharing rates in host communities, representing
the strength of potential interaction between viruses and any one host
(i.e., places where viruses undergo constant host switching may be
coevolutionary coldspots); (ii) the phylogenetic diversity of hosts, as
a proxy for variation in the immunological mechanisms that antagonize
viruses (i.e., the selection mosaic); and (iii) the local uniqueness of
the bat community, representing the potential for viruses to be exposed
to novel host traits (e.g., variation in receptor sequences). Together,
we argue that these can be used to identify and map the evolutionary
drivers that---in conjunction with transmission processes (e.g., viral
prevalence in reservoirs and animal-human contact rates)--- determine
disease emergence risk.

@@

Most previous work has assumed that the presence or richness of key
groups of bat hosts are predictive of coronavirus
diversity.\textsuperscript{2,3} Projecting bat and betacoronavirus
phylogeny over space (fig.~\ref{fig:biogeo}), we find support for the
idea that bat community assembly is directly responsible for a global
mosaic of viral evolution. The distinct groupings (represented by
different colors, symbolizing positions in a subspace formed by the
first two phylogenetic principal components) are essentially equivalent
between the two groups, and can be coarsely delineated as (1) south and
southeast Asia; (2) east Asia (including northern China), west Asia, and
the Mediterranean coast; (3) Eurasia above a northing of 40; and (4)
Africa and Latin America. In some cases, this diverges from expectations
about coronavirus biogeography: for example, previous work has rarely
flagged India as a region of interest, but for both bats and
betacoronaviruses, the subcontinent falls into the same regions as the
southeast Asian peninsula (and indeed, the region is home to known bat
hosts of multiple betacoronavirus subgenera, including nobecoviruses,
sarbecoviruses, and merbecoviruses).\textsuperscript{3}

@@

Overall, these results suggest that the boundaries of bat and
betacoronavirus biogeographic regions are largely consistent. This may
be surprising, given that cross-species transmission may play a stronger
role in coronavirus diversification than
cospeciation---\textsuperscript{2}a property that would theoretically
allow for substantial broad divergence in their biogeography. However,
host jumps at the family level or higher are relatively rare and
significant events in coronavirus evolutionary
history;\textsuperscript{2,14} as a result, the mosaic of
betacoronavirus phylogeography is assembled from a set of overlapping
smaller coevolutionary systems, superimposed in space and filtered by
the importance of different subgroups in local host communities. For
example, the most speciose and cosmopolitan family of bats, the vesper
bats (Vespertilionidae), are considered the primary hosts of the
subgenus \emph{Merbecovirus} (MERS-like viruses);\textsuperscript{3,14}
but in the Americas, where merbecoviruses are the only lineage present,
they have only been found in other bat taxa (\emph{e.g.}, Molossidae,
Phyllostomidae).\textsuperscript{15--18} At the coarsest scale, these
heterogeneities are lost, and betacoronavirus biogeography tracks the
deep rifts in bat evolutionary history---but within broad regions, the
component coevolutionary systems may have very different dynamics.

@@

Bats, the second most diverse groups of mammals, are found worldwide;
gradients in their species richness generally track broader patterns of
mammal diversity,\textsuperscript{\textbf{Tanalgo2022Mapping?}} with a
striking Neotropical hotspot (especially in the Amazon basin) and a
secondary hotspot centered in Indochina. These hotspots of bat diversity
are generally presumed to be hotspots of viral adaptive radiation, and
therefore areas of concern for human health.\textsuperscript{2,19}
However, the hotspots of known bat betacoronavirus hosts show a distinct
pattern, with primary hotspots (both in terms of area and higher values)
of host richness situated in southeast Asia, parts of southern Europe,
and to a lesser extent parts of Africa in the -25-0 range of latitudes
(fig.~\ref{fig:richness}; top). Although hundreds of species likely host
undiscovered betacoronaviruses, machine learning predictions have
suggested that these undiscovered reservoirs should follow the same
diversity gradient.\textsuperscript{20} In principle, these hotspots of
locally-diverse, virus-rich bat communities should drive more adaptive
diversification in their viruses.

@@

However, we find that the global pattern of betacoronavirus phylogenetic
distinctiveness is quite distinct from both bat host richness and
phylogenetic distinctiveness (fig.~\ref{fig:richness}; bottom). In
contrast to the sparsity of Neotropical betacoronavirus hosts, South and
Central America have the most evolutionary distinct hosts \emph{and}
viruses, followed by secondary hotspots in southeast Asia and the Rift
Valley region have mostly distinct viruses. Some degree of sampling bias
may contribute to these patterns: for example, the Neotropics are one of
the places where the fewest bat betacoronavirus sequences have been
generated,\textsuperscript{21--23} resulting in a sparser phylogenetic
tree, and artificially inflating distinctiveness; conversely,
disproportionate research effort in eastern China\textsuperscript{24}
may have led to a more complete inventory of the local diversity of
coronaviruses, again inflating these metrics relative to underlying
patterns. Even accounting for these potential biases, though, there is
obvious heterogeneity in betacoronavirus evolutionary distinctiveness
that is distinct from overall bat diversity.

Overall, these patterns recapitulate the evolutionary history of both
the order Chiroptera and the genus \emph{Betacoronavirus}. Horseshoe
bats (Rhinolophidae) include the reservoirs of the SARS-like viruses
(subgenus \emph{Sarbecovirus}), the group of pandemic threats that have
been of the greatest interest to researchers\textsuperscript{14} (and so
have been sampled most intensively).\textsuperscript{24} The hotspots of
host richness and viral diversity in southeast Asia---both of which are
disproportionately high, considering the global landscape of bat species
richness---are almost entirely driven by viral adaptive radiation
through host switching within this clade\textsuperscript{3,20}. In
contrast, the Neotropical hotspot of viral distinctiveness is driven by
isolation by host vicariance. Out of the four main groups of
betacoronaviruses, only merbecoviruses have been found in animals in the
Americas--- an introduction that is generally presumed to be
ancient.\textsuperscript{3,25} While comparatively understudied, New
World merbecoviruses have been found in the ghost-faced bats
(Mormoopidae), Neotropical leaf-nosed bats (Phyllostomidae), and
free-tailed bats (Molossidae).\textsuperscript{15--18} The former two
groups and a clade of the latter are endemic to the Neotropics, while
the explosive adaptive radiations of the phyllostomids are responsible
for the hotspot of bat diversity in the Amazon.\textsuperscript{26}
Together, these clades of New World bats play host to a distinct regime
of betacoronavirus coevolution.

@@

The existence of well-defined cophylogenetic regions suggests that the
bat-betacoronavirus system is spatially fragmented enough to create
divergent coevolutionary trajectories; in turn, this coevolutionary
mosaic may alter the risk of zoonotic emergence. These ideas are,
respectively, supported by the existence of hotspots of viral uniqueness
and the diverse origins of human betacoronaviruses. Together, this
framework points to a predictable relationship between host community
structure and coevolutionary pressure: phylogeographic structure in bat
hosts (and their diverse immune strategies)\textsuperscript{27} creates
a landscape of selective pressure; the trajectory of viruses'
coevolutionary response is, in turn, constrained by their opportunities
for either specialization or diversification through host jumps and
recombination.

@@

The greatest evolutionary potential for zoonotic emergence exists where
pathogen pools have a high genetic diversity and high propensity for
cross-species transmission. In our framework, emergence risk is
therefore maximized under higher phylogenetic diversity (viruses are
exposes to different host clades), higher host uniqueness (viruses are
experiencing novel, heterogeneous host traits combinations), and low to
medium viral sharing (host-virus pairs can coevolve independently, but
divergent viruses may still have opportunities for recombination). In
fig.~\ref{fig:trivariate}, this corresponds to yellow areas (dynamics
dominated by low viral sharing, with equal contributions of selection
mosaics and trait remixing; southeast Asia, and the Indian
sub-continent), green-yellow areas (dynamics with low viral sharing but
dominated by the selection mosaic effect of host diversity; sub-Saharan
Africa), and red-yellow areas (dynamics with low viral sharing but
dominated by trait remixing in host communities; the Middle East).
Translating this axis of variation back into a univariate risk map
(fig.~\ref{fig:risk}) highlights that this evolutionary landscape has a
striking correspondence to regions where zoonotic betacoronaviruses have
previously emerged.

Compared to approaches that map emergence risk based only on the number
of known bat hosts of betacoronaviruses, our framework suggests regions
where high viral sharing dominates coevolutionary dynamics---such as
Latin America, or Eurasia above a northing of 30---would pose less of a
relative risk of zoonotic emergence. Nevertheless, areas of high host
uniqueness coupled with high viral sharing (red-to-pink in
fig.~\ref{fig:trivariate}) could create hotspots facilitated by viral
codivergence. Our framework identifies Madagascar, where most bat
species are endemic following evolutionary divergence from sister
species in both African and Asian continents,\textsuperscript{28} as one
such hotspot; interestingly, a recent study\textsuperscript{29} reported
a novel and highly divergent lineage of nobecoviruses from
Madagascar-endemic pteropid bat species (\emph{Pteropus rufus} and
\emph{Rousettus madagascariensis}), again supporting the predictive
power of the coevolutionary framework.

@@

The relationship between the underlying pathogen pool and emergence risk
is mediated by both human-wildlife interfaces (the probability of
spillover) and opportunities for onward horizontal transmission (the
probability that spillovers become epidemics)\textsuperscript{1}. As a
proxy for both, we finally overlaid the risk component from the
composite map (see above) with the proportion of built land, as a proxy
for a mix of habitat disturbance, potential for bat synanthropy or
contact with bridge hosts like livestock,\textsuperscript{30,31} and
human population density and connectivity\textsuperscript{1,32,33}
(fig.~\ref{fig:compound}). Accounting for these factors, most of South
America and Europe are at comparatively lower risk, as--although densely
populated--settlements tend to be in areas with lower potential risk.
Conversely, regions like Malaysia and the northern coast of Australia
have a high evolutionary risk component, but should represent a
relatively lower effective risk due to low human density. However,
southeast Asia, the Indian subcontinent, and scattered hotspots in
sub-Saharan Africa are at high risk due to the overlap between human
populations and natural opportunities for cross-species transmission of
betacoronaviruses.

@@

Reassuringly, these predictions correspond to the geographic origins of
the three bat-origin coronaviruses that have recently emerged in human
populations. While available information puts the spillover of
SARS-CoV-2 in a live animal market in Wuhan, China, the ultimate origin
of the virus is almost certainly in a divergent lineage of
sarbecoviruses from Indochina that was poorly characterized prior to the
pandemic.\textsuperscript{21--23} Similarly, the SARS-CoV outbreak began
in Guangdong province in 2002, reaching humans through small carnivore
bridge hosts, but was eventually traced back to a set of likely
progenitor viruses found in cave-dwelling horseshoe bats in Yunnan
province;\textsuperscript{34} nearby, antibody evidence has indicated
human exposure to SARS-like viruses.\textsuperscript{35} MERS-CoV was
first detected in Jordan, but is widespread in camels in East Africa and
the Middle East, and may have reached its bridge host decades earlier
than originally supposed;\textsuperscript{36} as a result, the geography
of the original bat-to-camel transmission is still widely regarded as
uncertain. All of these are broadly consistent with the risk factors we
identify. Notably, India and west Africa are additional hotspots that
have yet to experience the emergence of a bat coronavirus into human
populations, but may still be at risk---particularly given known gaps in
bat surveillance,\textsuperscript{24} and a dense population in both
regions with global connectivity. In any of these regions, surveillance
on viral reservoirs can be paired with targeted monitoring of high-risk
human populations (i.e., those with regular wildlife
contact)\textsuperscript{37} for maximum impact.

@@

Bats emerged around 64 million years ago, and are one of the most
diverse mammalian orders, with more than 1,400 estimated
species.\textsuperscript{38,39} They exhibit a broad variety of habitat
use, behaviour, and feeding strategies, putting them at key positions in
the delivery and provisioning of several ecosystem services, tied to
important ecosystem-derived benefits to humans.\textsuperscript{40} Over
two-thirds of bats are know to be either obligate or facultative
insectivores, therefore actively contributing for agricultural pest
control,\textsuperscript{41,42} and vectors of pathogens that put a risk
on human health;\textsuperscript{43,44} some other species are essential
links in many seed-dispersal networks.\textsuperscript{45} However, many
of these species face a high risk of extinction, particularly given
persecution and killings that sometimes follows from messaging about
their role in disease emergence. Areas where bats, viruses, and humans
co-occur are not always hotspots of risk for human heath; as such,
developing more precise ways to map zoonotic hazards can help bats and
humans coexist safely, and support the conservation of these important
and unique animals.

>>>>

Disease emergence is complex, and is driven not only by animal-human
contact, but also by the underlying evolutionary dynamics in viral
reservoirs.\textsuperscript{\textbf{Plowright2017PatZoo?}} Although host
richness is often used as a superficial proxy for spillover
risk,\textsuperscript{\textbf{Anthony2017GloPat?},\textbf{Ruiz-Aravena2022EcoEvo?},\textbf{Sanchez2022Strategy?}}
these approaches oversimplify the relevant interspecific heterogeneity
in immunology, behavior, and other traits, and therefore overlook unique
host pools that allow for the rapid evolution of highly divergent
viruses.\textsuperscript{\textbf{Agosta2010HowSpe?}} In the case of
generalist pathogens like betacoronaviruses, there is conceptual and
empirical support to the idea that these community-level mechanisms are
even more important,\textsuperscript{\textbf{Power2004PatSpi?}}
particularly given that cross-species transmission may, as a rule,
structure viral evolution more than co-divergence with
hosts.\textsuperscript{\textbf{Geoghegan2017ComAna?}} This creates a
disconnect between coevolutionary theory and most existing ecological
frameworks for mapping spillover risk.

The geographic mosaic theory of coevolution (GMTC) attempts to
explicitly connect microevolutionary dynamics to the macroecology and
biogeography of symbiotic
interactions.\textsuperscript{\textbf{Thompson2005GeoMos?}} The GMTC
posits that coevolutionary processes among
pairs\textsuperscript{\textbf{Thompson1994CoePro?}} or
complexes\textsuperscript{\textbf{Janzen1980WheIt?}} of species are
structured in space by the rippling effects of abiotic conditions onto
evolutionary mechanisms, giving rise to fragmented systems with
different ecologies over large spatial
extents.\textsuperscript{\textbf{Price2002MacThe?}} The GMTC predicts a
spatial fragmentation of coevolutionary dynamics under the joint action
of three processes:\textsuperscript{\textbf{Gomulkiewicz2007DosDon?}}
coevolutionary hot- and coldspots, which appear when the intensity of
\emph{interaction} (in terms of reciprocal fitness consequences) varies
spatially; selection mosaics, wherein the intensity of \emph{selection}
varies across space, driven by both the biotic complexity of the
community (locally diverse hosts and viruses are more biotically
complex) and the local favorability of the
environment;\textsuperscript{\textbf{Thrall2007CoeSym?}} and trait
remixing, which occurs when coevolutionary dynamics change when
community-level \emph{functional traits} change through meta-community
dynamics.

Here, we apply the GMTC to explore and explain the global biogeography
of betacoronaviruses, the group that includes SARS-CoV, MERS-CoV, and
SARS-CoV-2. In their bat reservoirs, coronaviruses evolve through a mix
of host jumps, recombination among disparate lineages, and, to a lesser
degree, co-divergence with their
hosts---\textsuperscript{\textbf{Anthony2017GloPat?}}a mix of mechanisms
that creates a complex and nonlinear relationship between host diversity
and viral emergence. Working from a recently published database of bat
hosts of betacoronaviruses, we test whether spatial structure in
bat-betacoronavirus coevolution is identifiable at a global scale.
Aiming to explain these patterns, we develop a generalized framework for
applying the GMTC to host-virus interactions, with a specific emphasis
on the potential to create independent coevolutionary dynamics (and
therefore spatial fragmentation in risk) through heterogeneity. We
develop a trivariate risk assessment system that connects each GMTC
mechanism to a quantifiable aspect of host-virus interactions: (i) viral
sharing rates in host communities, representing the strength of
potential interaction between viruses and any one host (i.e., places
where viruses undergo constant host switching may be coevolutionary
coldspots); (ii) the phylogenetic diversity of hosts, as a proxy for
variation in the immunological mechanisms that antagonize viruses (i.e.,
the selection mosaic); and (iii) the local uniqueness of the bat
community, representing the potential for viruses to be exposed to novel
host traits (e.g., variation in receptor sequences). Together, we argue
that these can be used to identify and map the evolutionary drivers
that---in conjunction with transmission processes (e.g., viral
prevalence in reservoirs and animal-human contact rates)--- determine
disease emergence risk.

@@

Most previous work has assumed that the presence or richness of key
groups of bat hosts are predictive of coronavirus
diversity.\textsuperscript{\textbf{Anthony2017GloPat?},\textbf{Ruiz-Aravena2022EcoEvo?}}
Projecting bat and betacoronavirus phylogeny over space
(fig.~\ref{fig:biogeo}), we find support for the idea that bat community
assembly is directly responsible for a global mosaic of viral evolution.
The distinct groupings (represented by different colors, symbolizing
positions in a subspace formed by the first two phylogenetic principal
components) are essentially equivalent between the two groups, and can
be coarsely delineated as (1) south and southeast Asia; (2) east Asia
(including northern China), west Asia, and the Mediterranean coast; (3)
Eurasia above a northing of 40; and (4) Africa and Latin America. In
some cases, this diverges from expectations about coronavirus
biogeography: for example, previous work has rarely flagged India as a
region of interest, but for both bats and betacoronaviruses, the
subcontinent falls into the same regions as the southeast Asian
peninsula (and indeed, the region is home to known bat hosts of multiple
betacoronavirus subgenera, including nobecoviruses, sarbecoviruses, and
merbecoviruses).\textsuperscript{\textbf{Ruiz-Aravena2022EcoEvo?}}

@@

Overall, these results suggest that the boundaries of bat and
betacoronavirus biogeographic regions are broadly consistent at a global
scale; perfect matching between the biogeographic regions would have
indicated that the signal of virus distribution is fully predicted by
bat hosts ranges. Areas for which the biogeographic regions for bats and
betacoronaviruses differ are primarily (i) southeast Asia and southern
China, and (ii) the Arabian peninsula, which are both regions where
zoonotic transmission has been documented (potentially driving a unique
level of viral sampling effort that generates these patterns). These
spatially limited mismatches nonwithstanding, the large level of
congruence may be surprising, given that cross-species transmission may
play a stronger role in coronavirus diversification than
cospeciation---\textsuperscript{\textbf{Anthony2017GloPat?}}a property
that would theoretically allow for substantial broad divergence in their
biogeography. However, host jumps at the family level or higher are
relatively rare and significant events in coronavirus evolutionary
history;\textsuperscript{\textbf{Anthony2017GloPat?},\textbf{Latinne2020OriCro?}}
as a result, the mosaic of betacoronavirus phylogeography is assembled
from a set of overlapping smaller coevolutionary systems, superimposed
in space and filtered by the importance of different subgroups in local
host communities. For example, the most speciose and cosmopolitan family
of bats, the vesper bats (Vespertilionidae), are considered the primary
hosts of the subgenus \emph{Merbecovirus} (MERS-like
viruses);\textsuperscript{\textbf{Latinne2020OriCro?},\textbf{Ruiz-Aravena2022EcoEvo?}}
but in the Americas, where merbecoviruses are the only lineage present,
they have only been found in other bat taxa (\emph{e.g.}, Molossidae,
Phyllostomidae).\textsuperscript{\textbf{Anthony2013CorBat?},\textbf{Goes2013NovBat?},\textbf{Goes2016GenDiv?},\textbf{Brandao2008CorDet?}}
At the coarsest scale, these heterogeneities are lost, and
betacoronavirus biogeography tracks the deep rifts in bat evolutionary
history---but within broad regions, the component coevolutionary systems
may have very different dynamics.

@@

Bats, the second most diverse groups of mammals, are found worldwide;
gradients in their species richness generally track broader patterns of
mammal diversity,\textsuperscript{\textbf{Tanalgo2022Mapping?}} with a
striking Neotropical hotspot (especially in the Amazon basin) and a
secondary hotspot centered in Indochina. These hotspots of bat diversity
are generally presumed to be hotspots of viral adaptive radiation, and
therefore areas of concern for human
health.\textsuperscript{\textbf{Anthony2017GloPat?},\textbf{Olival2017HosVir?}}
However, the hotspots of known bat betacoronavirus hosts show a distinct
pattern, with primary hotspots (both in terms of area and higher values)
of host richness situated in southeast Asia, parts of southern Europe,
and to a lesser extent parts of Africa in the -25-0 range of latitudes
(fig.~\ref{fig:richness}; top). Although hundreds of species likely host
undiscovered betacoronaviruses, machine learning predictions have
suggested that these undiscovered reservoirs should follow the same
diversity gradient.\textsuperscript{\textbf{Becker2022OptPre?}} In
principle, these hotspots of locally-diverse, virus-rich bat communities
should drive more adaptive diversification in their viruses.

@@

However, we find that the global pattern of betacoronavirus phylogenetic
distinctiveness is quite distinct from both bat host richness and
phylogenetic distinctiveness (fig.~\ref{fig:richness}; bottom). In
contrast to the sparsity of Neotropical betacoronavirus hosts, South and
Central America have the most evolutionary distinct hosts \emph{and}
viruses, followed by secondary hotspots in southeast Asia and the Rift
Valley region have mostly distinct viruses. Some degree of sampling bias
may contribute to these patterns: for example, the Neotropics are one of
the places where the fewest bat betacoronavirus sequences have been
generated,\textsuperscript{\textbf{Worobey2022HuaMar?},\textbf{Temmam2022BatCor?},\textbf{Boni2020EvoOri?}}
resulting in a sparser phylogenetic tree, and artificially inflating
distinctiveness; conversely, disproportionate research effort in eastern
China\textsuperscript{\textbf{Cohen2022SamStr?}} may have led to a more
complete inventory of the local diversity of coronaviruses, again
inflating these metrics relative to underlying patterns. Even accounting
for these potential biases, though, there is obvious heterogeneity in
betacoronavirus evolutionary distinctiveness that is distinct from
overall bat diversity.

Overall, these patterns recapitulate the evolutionary history of both
the order Chiroptera and the genus \emph{Betacoronavirus}. Horseshoe
bats (Rhinolophidae) include the reservoirs of the SARS-like viruses
(subgenus \emph{Sarbecovirus}), the group of pandemic threats that have
been of the greatest interest to
researchers\textsuperscript{\textbf{Latinne2020OriCro?}} (and so have
been sampled most
intensively).\textsuperscript{\textbf{Cohen2022SamStr?}} The hotspots of
host richness and viral diversity in southeast Asia---both of which are
disproportionately high, considering the global landscape of bat species
richness---are almost entirely driven by viral adaptive radiation
through host switching within this
clade\textsuperscript{\textbf{Becker2022OptPre?},\textbf{Ruiz-Aravena2022EcoEvo?}}.
In contrast, the Neotropical hotspot of viral distinctiveness is driven
by isolation by host vicariance. Out of the four main groups of
betacoronaviruses, only merbecoviruses have been found in animals in the
Americas--- an introduction that is generally presumed to be
ancient.\textsuperscript{\textbf{Ruiz-Aravena2022EcoEvo?},\textbf{Olival2020PosRev?}}
While comparatively understudied, New World merbecoviruses have been
found in the ghost-faced bats (Mormoopidae), Neotropical leaf-nosed bats
(Phyllostomidae), and free-tailed bats
(Molossidae).\textsuperscript{\textbf{Anthony2013CorBat?},\textbf{Goes2013NovBat?},\textbf{Goes2016GenDiv?},\textbf{Brandao2008CorDet?}}
The former two groups and a clade of the latter are endemic to the
Neotropics, while the explosive adaptive radiations of the phyllostomids
are responsible for the hotspot of bat diversity in the
Amazon.\textsuperscript{\textbf{Ammerman2012FirMol?}} Together, these
clades of New World bats play host to a distinct regime of
betacoronavirus coevolution.

Our approach is potentially limited by sampling bias: key hotspots
identified by our model have, indeed, been sampled intensely following
major zoonotic emergence events. In these areas, more betacoronavirus
hosts will have been discovered, leading to higher overall diversity and
potentially higher sharing. Similarly, hotspots of evolutionary
uniqueness - as in the Arabian peninsula - could reflect much broader
lineages that have only been sampled in focal areas for public health.
While the discovery of new branches of bat-betacoronavirus coevolution
is certainly likely, and might change some of the observed patterns, our
framework is likely to be fairly robust: the 126 hosts in our study
capture nearly 10\% of global bat diversity, and the underlying
evolutionary patterns they represent are much less sensitive to new
information than any inferences about viral evolution.

@@

The existence of well-defined cophylogenetic regions suggests that the
bat-betacoronavirus system is spatially fragmented enough to create
divergent coevolutionary trajectories; in turn, this coevolutionary
mosaic may alter the risk of zoonotic emergence. These ideas are,
respectively, supported by the existence of hotspots of viral uniqueness
and the diverse origins of human betacoronaviruses. Together, this
framework points to a predictable relationship between host community
structure and coevolutionary pressure: phylogeographic structure in bat
hosts (and their diverse immune
strategies)\textsuperscript{\textbf{Banerjee2020NovIns?}} creates a
landscape of selective pressure; the trajectory of viruses'
coevolutionary response is, in turn, constrained by their opportunities
for either specialization or diversification through host jumps and
recombination.

@@

The greatest evolutionary potential for zoonotic emergence exists where
pathogen pools have a high genetic diversity and high propensity for
cross-species transmission. In our framework, emergence risk is
therefore maximized under higher phylogenetic diversity (viruses are
exposed to different host clades), higher host uniqueness (viruses are
experiencing novel, heterogeneous host traits combinations), and low to
medium viral sharing (host-virus pairs can coevolve independently, but
divergent viruses may still have opportunities for recombination). In
fig.~\ref{fig:trivariate}, this corresponds to yellow areas (dynamics
dominated by low viral sharing, with equal contributions of selection
mosaics and trait remixing; southeast Asia, and the Indian
sub-continent), green-yellow areas (dynamics with low viral sharing but
dominated by the selection mosaic effect of host diversity; sub-Saharan
Africa), and red-yellow areas (dynamics with low viral sharing but
dominated by trait remixing in host communities; the Middle East).
Translating this axis of variation back into a univariate risk map
(fig.~\ref{fig:risk}) highlights that this evolutionary landscape has a
striking correspondence to regions where zoonotic betacoronaviruses have
previously emerged. Our findings align with predictions regarding the
spatial location of cross-species transmission. These locations not only
pose a potential risk of viral jumps that could endanger human health
but also provide valuable information for monitoring wildlife health.
This could guide us to determine where and what measures to implement
for effectively monitoring wildlife and human betacoronavirus outbreaks
before they escalate to critical levels. Nevertheless, there are
actually very few documented cases of emergence events, and similarities
could be some degree of coincidental.

Compared to approaches that map emergence risk based only on the number
of known bat hosts of betacoronaviruses, our framework suggests regions
where high viral sharing dominates coevolutionary dynamics---such as
Latin America, or Eurasia above a northing of 30---would pose less of a
relative risk of zoonotic emergence. Nevertheless, areas of high host
uniqueness coupled with high viral sharing (red-to-pink in
fig.~\ref{fig:trivariate}) could create hotspots facilitated by viral
codivergence. Our framework identifies Madagascar, where most bat
species are endemic following evolutionary divergence from sister
species in both African and Asian
continents,\textsuperscript{\textbf{Shi2014DeeDiv?}} as one such
hotspot; interestingly, a recent
study\textsuperscript{\textbf{Kettenburg2022FulGen?}} reported a novel
and highly divergent lineage of nobecoviruses from Madagascar-endemic
pteropid bat species (\emph{Pteropus rufus} and \emph{Rousettus
madagascariensis}), again supporting the predictive power of the
coevolutionary framework.

@@

The relationship between the underlying pathogen pool and emergence risk
is mediated by both human-wildlife interfaces (the probability of
spillover) and opportunities for onward horizontal transmission (the
probability that spillovers become
epidemics)\textsuperscript{\textbf{Plowright2017PatZoo?}}. It must be
noted that the assesment of risk based on the GMTC mechanisms does not
account for human presence; for this reason, it represents ``potential''
level of risk, which must be re-evaluated in the light of human
presence. As a proxy for both, we finally overlaid the risk component
from the composite map (see above) with the proportion of built land, as
a proxy for a mix of habitat disturbance, potential for bat synanthropy
or contact with bridge hosts like
livestock,\textsuperscript{\textbf{Rulli2021LanCha?},\textbf{Cui2019OriEvo?}}
and human population density and
connectivity\textsuperscript{\textbf{Plowright2017PatZoo?},\textbf{Muylaert2022PreFut?},\textbf{Hassell2017UrbDis?}}
(fig.~\ref{fig:compound}). Accounting for these factors, most of South
America and Europe are at comparatively lower risk, as--although densely
populated--settlements tend to be in areas with lower potential risk.
Conversely, regions like Malaysia and the northern coast of Australia
have a high evolutionary risk component, but should represent a
relatively lower effective risk due to low human density. However,
southeast Asia, the Indian subcontinent, and scattered hotspots in
sub-Saharan Africa are at high risk due to the overlap between human
populations and natural opportunities for cross-species transmission of
betacoronaviruses.

@@

Reassuringly, these predictions correspond to the geographic origins of
the three bat-origin coronaviruses that have recently emerged in human
populations. While available information puts the spillover of
SARS-CoV-2 in a live animal market in Wuhan, China, the ultimate origin
of the virus is almost certainly in a divergent lineage of
sarbecoviruses from Indochina that was poorly characterized prior to the
pandemic.\textsuperscript{\textbf{Worobey2022HuaMar?},\textbf{Temmam2022BatCor?},\textbf{Boni2020EvoOri?}}
Similarly, the SARS-CoV outbreak began in Guangdong province in 2002,
reaching humans through small carnivore bridge hosts, but was eventually
traced back to a set of likely progenitor viruses found in cave-dwelling
horseshoe bats in Yunnan
province;\textsuperscript{\textbf{Hu2017DisRic?}} nearby, antibody
evidence has indicated human exposure to SARS-like
viruses.\textsuperscript{\textbf{Wang2018SerEvi?}} MERS-CoV was first
detected in Jordan, but is widespread in camels in East Africa and the
Middle East, and may have reached its bridge host decades earlier than
originally supposed;\textsuperscript{\textbf{Muller2014MerCor?}} as a
result, the geography of the original bat-to-camel transmission is still
widely regarded as uncertain. All of these are broadly consistent with
the risk factors we identify. Notably, India and west Africa are
additional hotspots that have yet to experience the emergence of a bat
coronavirus into human populations, but may still be at
risk---particularly given known gaps in bat
surveillance,\textsuperscript{\textbf{Cohen2022SamStr?}} and a dense
population in both regions with global connectivity. In any of these
regions, surveillance on viral reservoirs can be paired with targeted
monitoring of high-risk human populations (i.e., those with regular
wildlife contact)\textsuperscript{\textbf{Xu2004EpiClu?}} for maximum
impact.

@@

Bats emerged around 64 million years ago, and are one of the most
diverse mammalian orders, with more than 1,400 estimated
species.\textsuperscript{\textbf{Peixoto2018SynEco?},\textbf{Simmons2020BatSpe?}}
They exhibit a broad variety of habitat use, behaviour, and feeding
strategies, putting them at key positions in the delivery and
provisioning of several ecosystem services, tied to important
ecosystem-derived benefits to
humans.\textsuperscript{\textbf{Kasso2013EcoEco?}} Over two-thirds of
bats are known to be either obligate or facultative insectivores,
therefore actively contributing for agricultural pest
control,\textsuperscript{\textbf{Voigt2016BatAnt?},\textbf{Williams-Guillen2008BatLim?}}
and vectors of pathogens that put a risk on human
health;\textsuperscript{\textbf{Gonsalves2013MosCon?},\textbf{Gonsalves2013MosInf?}}
some other species are essential links in many seed-dispersal
networks.\textsuperscript{\textbf{Mello2011MisPar?}} However, many of
these species face a high risk of extinction, particularly given
persecution and killings that sometimes follows from messaging about
their role in disease emergence. Areas where bats, viruses, and humans
co-occur are not always hotspots of risk for human heath; as such,
developing more precise ways to map zoonotic hazards can help bats and
humans coexist safely, and support the conservation of these important
and unique animals.
